---
title: "Epi R Handbook Notes"
output: html_document
date: "2024-07-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Management

## Cleaning data and core functions 0%

## Working with dates 0%

## Characters and strings 0%

## Factors 0%

## Pivoting data 0%

## Grouping data 0%

## Joining data 0%

## De-duplication 0%

## Iteration, loops and lists 0%

# Analysis

## Descriptive tables 0%

## Simple statistical tests 0%

## Univariate and multivariable regression 0%

## Missing data 0%

## Standardised rates 0%

## Moving averages 0%

## Time series and outbreak detection 0%

## Epidemic modelling 0%

## Contact tracing 0%

## Survey analysis 0%

```{r}
## load packages from CRAN
pacman::p_load(rio,          # File import
               here,         # File locator
               tidyverse,    # data management + ggplot2 graphics
               tsibble,      # handle time series datasets
               survey,       # for survey functions
               srvyr,        # dplyr wrapper for survey package
               gtsummary,    # wrapper for survey package to produce tables
               apyramid,     # a package dedicated to creating age pyramids
               patchwork,    # for combining ggplots
               ggforce       # for alluvial/sankey plots
               ) 

## load packages from github
pacman::p_load_gh(
     "R4EPI/sitrep"          # for observation time / weighting functions
)
```

```{r}
# import the survey data
survey_data <- import(here("data", "survey_data.xlsx"))

# import the dictionary into R
survey_dict <- import(here("data", "survey_dict.xlsx"))

# import the population data
population <- import(here("data", "population.xlsx"))
```

```{r}
## select the date variable names from the dictionary 
DATEVARS <- survey_dict %>% 
  filter(type == "date") %>% 
  filter(name %in% names(survey_data)) %>% 
  ## filter to match the column names of your data
  pull(name) # select date vars
  
## change to dates 
survey_data <- survey_data %>%
  mutate(across(all_of(DATEVARS), as.Date))


## add those with only age in months to the year variable (divide by twelve)
survey_data <- survey_data %>% 
  mutate(age_years = if_else(is.na(age_years), 
                             age_months / 12, 
                             age_years))

## define age group variable
survey_data <- survey_data %>% 
     mutate(age_group = age_categories(age_years, 
                                    breakers = c(0, 3, 15, 30, 45)
                                    ))


## create a character variable based off groups of a different variable 
survey_data <- survey_data %>% 
  mutate(health_district = case_when(
    cluster_number %in% c(1:5) ~ "district_a", 
    TRUE ~ "district_b"
  ))

## select the yes/no variable names from the dictionary 
YNVARS <- survey_dict %>% 
  filter(type == "yn") %>% 
  filter(name %in% names(survey_data)) %>% 
  ## filter to match the column names of your data
  pull(name) # select yn vars
  
## change to dates 
survey_data <- survey_data %>%
  mutate(across(all_of(YNVARS), 
                str_detect, 
                pattern = "yes"))
```

```{r}
# ## join the individual and household data to form a complete data set
# survey_data <- left_join(survey_data_hh, 
#                          survey_data_indiv,
#                          by = c("_index" = "_parent_index"))
# 
# 
# ## create a unique identifier by combining indeces of the two levels 
# survey_data <- survey_data %>% 
#      mutate(uid = str_glue("{index}_{index_y}"))


#Note: this as already been done in the dataset. Not sure what the indices refer to.
```

```{r}
## set the start/end of recall period
## can be changed to date variables from dataset 
## (e.g. arrival date & date questionnaire)
survey_data <- survey_data %>% 
  mutate(recall_start = as.Date("2018-01-01"), 
         recall_end = as.Date("2018-05-01"))


# set inappropriate dates to NA based on rules 
## e.g. arrivals before start, departures departures after end
survey_data <- survey_data %>%
      mutate(
           arrived_date = if_else(arrived_date < recall_start, 
                                 as.Date(NA),
                                  arrived_date),
           birthday_date = if_else(birthday_date < recall_start,
                                  as.Date(NA),
                                  birthday_date),
           left_date = if_else(left_date > recall_end,
                              as.Date(NA),
                               left_date),
           death_date = if_else(death_date > recall_end,
                               as.Date(NA),
                               death_date)
           )
```

```{r}
## create new variables for start and end dates/causes
survey_data <- survey_data %>% 
     ## choose earliest date entered in survey
     ## from births, household arrivals, and camp arrivals 
     find_start_date("birthday_date",
                  "arrived_date",
                  period_start = "recall_start",
                  period_end   = "recall_end",
                  datecol      = "startdate",
                  datereason   = "startcause" 
                 ) %>%
     ## choose earliest date entered in survey
     ## from camp departures, death and end of the study
     find_end_date("left_date",
                "death_date",
                period_start = "recall_start",
                period_end   = "recall_end",
                datecol      = "enddate",
                datereason   = "endcause" 
               )


## label those that were present at the start/end (except births/deaths)
survey_data <- survey_data %>% 
     mutate(
       ## fill in start date to be the beginning of recall period (for those empty) 
       startdate = if_else(is.na(startdate), recall_start, startdate), 
       ## set the start cause to present at start if equal to recall period 
       ## unless it is equal to the birth date 
       startcause = if_else(startdate == recall_start & startcause != "birthday_date",
                              "Present at start", startcause), 
       ## fill in end date to be end of recall period (for those empty) 
       enddate = if_else(is.na(enddate), recall_end, enddate), 
       ## set the end cause to present at end if equall to recall end 
       ## unless it is equal to the death date
       endcause = if_else(enddate == recall_end & endcause != "death_date", 
                            "Present at end", endcause))


## Define observation time in days
survey_data <- survey_data %>% 
  mutate(obstime = as.numeric(enddate - startdate))
```


## Survival analysis 0%

## GIS basics 0%

# Data visualisation

## Tables for presentation 0%

## ggplot basics 0%

## ggplot tips 0%

## Epidemic curves 0%

## Demographic pyramids and Likert scales 0%

## Heat plots 0%

## Diagrams and charts 0%

## Combinations analysis 0%

## Transmission chains 0%

## Phylogenetic trees 0%

## Interactive plots 0%
