---
title: "Epi R Handbook Notes"
output: html_document
date: "2024-07-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Management

## Cleaning data and core functions 0%

## Working with dates 0%

## Characters and strings 0%

## Factors 0%

## Pivoting data 0%

## Grouping data 0%

## Joining data 0%

## De-duplication 0%

## Iteration, loops and lists 0%

# Analysis

## Descriptive tables 0%

## Simple statistical tests 0%

## Univariate and multivariable regression 0%

## Missing data 0%

## Standardised rates 0%

## Moving averages 0%

## Time series and outbreak detection 0%

## Epidemic modelling 0%

## Contact tracing 0%

## Survey analysis 0%

```{r}
## load packages from CRAN
pacman::p_load(rio,          # File import
               here,         # File locator
               tidyverse,    # data management + ggplot2 graphics
               tsibble,      # handle time series datasets
               survey,       # for survey functions
               srvyr,        # dplyr wrapper for survey package
               gtsummary,    # wrapper for survey package to produce tables
               apyramid,     # a package dedicated to creating age pyramids
               patchwork,    # for combining ggplots
               ggforce       # for alluvial/sankey plots
               ) 

## load packages from github
pacman::p_load_gh(
     "R4EPI/sitrep"          # for observation time / weighting functions
)
```

## Survival analysis 0%

## GIS basics 0%

# Data visualisation

## Tables for presentation 0%

## ggplot basics 0%

## ggplot tips 0%

## Epidemic curves 0%

## Demographic pyramids and Likert scales 0%

## Heat plots 0%

## Diagrams and charts 0%

## Combinations analysis 0%

## Transmission chains 0%

## Phylogenetic trees 0%

## Interactive plots 0%
